% -------------------------------------------------------
% SECTION 3: THE QUANTUM PROPHETIC WORLD
% -------------------------------------------------------

We now introduce the formal structure of a \emph{quantum prophetic world}. Unlike classical
predictive systems, a prophetic world contains an explicit feedback loop: the act of revealing
information about future states influences the evolution of those states. Our construction begins
with the specification of the Hilbert space, the initial joint state, and the world dynamics.

\subsection{State Space}

Let $\mathcal{H} = \mathcal{H}_X \otimes \mathcal{H}_V$ where:
\begin{itemize}[leftmargin=1.0cm]
    \item $\mathcal{H}_X$ is the latent (future) variable register, typically encoding future
    environmental states such as crash or non-crash conditions;
    \item $\mathcal{H}_V$ is the prophecy register, entangled with $\mathcal{H}_X$ and accessible
    to measurement by an observer or regulator.
\end{itemize}

For the concrete instantiation used throughout this monograph, each subsystem is a qubit,
$\mathcal{H}_X \cong \mathbb{C}^2$ and $\mathcal{H}_V \cong \mathbb{C}^2$, although the following
formalism extends to arbitrary finite dimensions.

\subsection{Initial Joint State}

\begin{definition}[Initial Prophetic State]
An initial quantum prophetic world is given by a state $\rho_0 \in \mathcal{D}(\mathcal{H})$.
For concreteness, we consider the one-parameter family
\[
    \rho_0 = \dyad{\psi_0},
    \qquad
    \ket{\psi_0}
    = \sqrt{p}\,\ket{00} + \sqrt{1-p}\,\ket{11},
\]
with $p \in [0,1]$.
\end{definition}

This state has the property that the marginal of the prophecy register $\mathcal{H}_V$ contains
information about the future latent state encoded in $\mathcal{H}_X$, but not perfectly so. Such
partial predictability is essential for non-trivial prophetic behavior.

\subsection{World Dynamics as a CPTP Evolution}

The environment evolves according to a CPTP map $\mathcal{E}$ acting non-trivially on the latent
register $\mathcal{H}_X$ and trivially on the prophecy register.

\begin{definition}[World Evolution]
A quantum prophetic world evolves under a map
\[
    \mathcal{E} : \mathcal{D}(\mathcal{H}) \to \mathcal{D}(\mathcal{H}),
\]
given in Kraus form by
\[
    \mathcal{E}(\rho)
    =
    \sum_{i}
    (K_i \otimes \id_V) \, \rho \,
    (K_i^\dagger \otimes \id_V),
\]
where $\{K_i\}$ are Kraus operators on $\mathcal{H}_X$ satisfying
$\sum_i K_i^\dagger K_i = \id_X$.
\end{definition}

For explicitness, we adopt the amplitude-damping channel:
\[
    K_0 =
    \begin{pmatrix}
        1 & 0 \\
        0 & \sqrt{1-\gamma}
    \end{pmatrix},
    \qquad
    K_1 =
    \begin{pmatrix}
        0 & \sqrt{\gamma} \\
        0 & 0
    \end{pmatrix},
\]
where $\gamma \in [0,1]$ is a crash-probability amplification parameter.

\begin{remark}
Amplitude damping introduces asymmetry between the $\ket{0}$ and $\ket{1}$ states of the latent
register, matching the intuition that “crash-like” failures may increase over time.
\end{remark}

\subsection{Raw Prophecy Extraction}

\begin{definition}[Raw Prophecy Channel]
The raw prophecy accessible to a measurement device is the marginal
\[
    \Phi_{\mathrm{raw}}(\rho) = \Tr_X(\rho) \in \mathcal{D}(\mathcal{H}_V).
\]
\end{definition}

This captures the idea that prophecy consists of a (possibly lossy) viewpoint onto future states.

\medskip
This completes the formal definition of the quantum prophetic world. We now describe the
measurement, regulatory, and agent-response mechanisms that interact with it.

% -------------------------------------------------------
% SECTION 4: THE PROPHECY–RESPONSE OPERATOR \Phi^{\mathcal{Q}}
% -------------------------------------------------------
\section{The Prophecy--Response Operator $\Phi^{\mathcal{Q}}$}

We now formalize the fundamental operator governing the evolution of a quantum prophetic world
when prophecy is observed, regulated, acted upon, and then propagated forward. The operator
$\Phi^{\mathcal{Q}}$ integrates four components:
\[
    \Phi^{\mathcal{Q}}
    =
    \mathcal{E}
    \circ
    \pi
    \circ
    \rho_{\mathrm{rel}}
    \circ
    M.
\]

\subsection{Prophetic Measurement}

Let $\{E_i\}$ be a binary POVM on $\mathcal{H}_V$ with sharpness parameter $\alpha$:
\[
    E_0 = (1-\alpha)\frac{\id}{2} + \alpha \dyad{0},
    \qquad
    E_1 = (1-\alpha)\frac{\id}{2} + \alpha \dyad{1}.
\]

\begin{definition}[Measurement Instrument]
Define $M_i = \sqrt{E_i}$ and
\[
    M(\rho)
    =
    \sum_{i=0}^1
    (\id_X \otimes M_i)\, \rho\,
    (\id_X \otimes M_i^\dagger).
\]
\end{definition}

By construction, $M$ is CPTP and physically implementable.

\subsection{Regulated Prophecy Release}

The regulator transforms the post-measurement reduced prophecy. We use a depolarizing channel.

\begin{definition}[Regulator Channel]
For $\lambda \in [0,1]$, define
\[
    \rho_{\mathrm{rel}}(\sigma)
    =
    (1-\lambda)\, \sigma + \lambda\, \frac{\id}{2}.
\]
This induces controlled information leakage.
\end{definition}

\subsection{Agent Response}

Agents react to the released prophecy via conditional unitary updates.

\begin{definition}[Agent Response Map]
Let $U_0 = \id_X$ and $U_1 = X$ (Pauli-$X$ on $\mathcal{H}_X$). Define
\[
    \pi(\rho)
    =
    \sum_{i=0}^1
    (U_i \otimes \id_V)\,
    (\id_X \otimes M_i) \rho (\id_X \otimes M_i^\dagger)\,
    (U_i^\dagger \otimes \id_V).
\]
\end{definition}

This reflects a minimal model where the agent intervenes on the latent quantum state when prophecy
indicates a potential crash.

\subsection{Definition of the Combined Operator}

\begin{definition}[Quantum Prophecy--Response Operator]
The composed operator $\Phi^{\mathcal{Q}} : \mathcal{D}(\mathcal{H}) \to \mathcal{D}(\mathcal{H})$ is
\[
    \Phi^{\mathcal{Q}}(\rho)
    =
    \mathcal{E}\!\left(
    \pi\!\left(
        \rho_{\mathrm{rel}}(
            M(\rho)
        )
    \right)
    \right).
\]
\end{definition}

We expand the full expression explicitly using Kraus operators.

\begin{proposition}[Explicit Expansion]
Let $\{K_j\}$ be the Kraus operators of $\mathcal{E}$. Then
\[
\begin{aligned}
    \Phi^{\mathcal{Q}}(\rho)
    &=
    \sum_{j=0}^1
    \sum_{i=0}^1
    (K_j U_i \otimes M_i)
    \Big[
        (1-\lambda)\rho + \lambda\, \tfrac{\id}{4}
    \Big]
    (U_i^\dagger K_j^\dagger \otimes M_i^\dagger).
\end{aligned}
\]
\end{proposition}

\begin{proof}
Each component map ($M$, $\rho_{\mathrm{rel}}$, $\pi$, $\mathcal{E}$) admits a Kraus representation.
The sequential composition of CPTP maps is CPTP, and the Kraus operators multiply accordingly.
\end{proof}

\subsection{Basic Properties}

\begin{theorem}[$\Phi^{\mathcal{Q}}$ is CPTP]
The operator $\Phi^{\mathcal{Q}}$ is completely positive and trace preserving.
\end{theorem}

\begin{proof}
Each constituent map is CPTP:
\begin{itemize}
    \item $M$ is CPTP as a measurement instrument;
    \item $\rho_{\mathrm{rel}}$ is CPTP as a depolarizing channel;
    \item $\pi$ is CPTP as a conditional unitary channel;
    \item $\mathcal{E}$ is CPTP by definition.
\end{itemize}
The composition of CPTP maps is CPTP.
\end{proof}

\begin{theorem}[Sufficient Conditions for Contraction]
If $0 < \lambda \le \lambda_0$, $0 \le \alpha \le \alpha_0$, and $0 \le \gamma \le \gamma_0$ for sufficiently
small constants, then
\[
    \|\Phi^{\mathcal{Q}}(\rho_1) - \Phi^{\mathcal{Q}}(\rho_2)\|_1
    \le k
    \|\rho_1 - \rho_2\|_1
\]
for some $k < 1$.
\end{theorem}

\begin{proof}[Sketch]
Each component map has Lipschitz constant at most~$1$. The depolarizing channel introduces
strict contraction in trace norm; the remaining maps can be made arbitrarily close to isometries
by taking $\alpha$ and $\gamma$ sufficiently small. The product of their Lipschitz constants is then
less than~$1$.
\end{proof}

\begin{corollary}[Existence of Fixed Point]
Under the above conditions, $\Phi^{\mathcal{Q}}$ admits a unique fixed point $\rho^\ast$.
\end{corollary}

This concludes the formal specification of the prophecy--response operator. Ethical constraints
will be integrated in Section~5.
