% -------------------------------------------------------
% SECTION 8: WORKED EXAMPLE AND IMPLEMENTATION
% -------------------------------------------------------

This section presents a full-scale worked example of a quantum prophetic world and a concrete
implementation recipe for the prophecy--response operator $\Phi^{\mathcal{Q}}$. We give explicit
matrices, compute a complete iteration of $\Phi^{\mathcal{Q}}$ on a chosen initial state, show a
numerical contraction test, evaluate ethical metrics (harm and leakage), and demonstrate a
simple optimization (grid search) for regulator parameters. The example is fully reproducible and
serves as a blueprint for simulation using standard quantum libraries (QuTiP, Qiskit, or NumPy).

\subsection{Example Parameterization}

We instantiate the minimal two-qubit system with the following parameter choices (numerical
values used throughout):

\begin{align*}
    p &= 0.7 &\text{(initial entanglement parameter)}\\
    \gamma &= 0.2 &\text{(amplitude-damping parameter for the latent register)}\\
    \alpha &= 0.6 &\text{(measurement sharpness)}\\
    \lambda &= 0.3 &\text{(regulator depolarizing strength)}.
\end{align*}

All matrices are given in the computational basis ordered as $\{\ket{00},\ket{01},\ket{10},\ket{11}\}$
with the first qubit corresponding to the latent register $X$ and the second to the prophecy register
$V$.

\subsection{Initial State}

Define the initial pure state
\[
    \ket{\psi_0} = \sqrt{p}\,\ket{00} + \sqrt{1-p}\,\ket{11},
    \qquad
    \rho_0 = \ket{\psi_0}\bra{\psi_0}.
\]

For $p=0.7$, the numeric density matrix $\rho_0$ is
\[
\rho_0 =
\begin{pmatrix}
0.700000 & 0       & 0       & 0.458258\\[4pt]
0        & 0       & 0       & 0       \\[4pt]
0        & 0       & 0       & 0       \\[4pt]
0.458258 & 0       & 0       & 0.300000
\end{pmatrix},
\]
where we show entries to six decimal places.

\subsection{World Evolution: Amplitude Damping on $X$}

We take the amplitude-damping Kraus operators on $\mathcal{H}_X$,
\[
K_0=
\begin{pmatrix}
1 & 0\\[3pt]
0 & \sqrt{1-\gamma}
\end{pmatrix},\qquad
K_1=
\begin{pmatrix}
0 & \sqrt{\gamma}\\[3pt]
0 & 0
\end{pmatrix}.
\]
With $\gamma=0.2$ these evaluate to
\[
K_0=\begin{pmatrix}1 & 0\\[3pt] 0 & 0.894427\end{pmatrix},\qquad
K_1=\begin{pmatrix}0 & 0.447214\\[3pt] 0 & 0\end{pmatrix}.
\]
On the full two-qubit space these act as $(K_j\otimes I_V)$ in the operator-sum for $\mathcal{E}$.

\subsection{Prophetic Measurement: POVM and Kraus (Instrument) Operators}

We choose the binary POVM on $\mathcal{H}_V$ parameterized by $\alpha$:
\[
E_0 = (1-\alpha)\tfrac{\id}{2} + \alpha \ket{0}\!\bra{0},\qquad
E_1 = (1-\alpha)\tfrac{\id}{2} + \alpha \ket{1}\!\bra{1}.
\]
For $\alpha = 0.6$ the POVM elements are diagonal in the computational basis with eigenvalues
\[
E_0 = \mathrm{diag}\!\big(\tfrac{1+\alpha}{2},\,\tfrac{1-\alpha}{2}\big)
     = \mathrm{diag}\!\big(0.8,\,0.2\big),
\]
\[
E_1 = \mathrm{diag}\!\big(0.2,\,0.8\big).
\]

We take the instrument with Kraus operators $M_i=\sqrt{E_i}$ (which are diagonal here):
\[
M_0=\mathrm{diag}\!\left(\sqrt{0.8},\sqrt{0.2}\right)
    \approx \begin{pmatrix}0.894427 & 0\\[3pt] 0 & 0.447214\end{pmatrix},
\]
\[
M_1=\mathrm{diag}\!\left(\sqrt{0.2},\sqrt{0.8}\right)
    \approx \begin{pmatrix}0.447214 & 0\\[3pt] 0 & 0.894427\end{pmatrix}.
\]

These produce the measurement instrument
\[
M(\rho) = \sum_{i=0}^1 (\id_X\otimes M_i)\;\rho\;(\id_X\otimes M_i^\dagger),
\]
which is CPTP.

\subsection{Regulator Channel: Depolarizing on $V$}

The regulator applies a depolarizing channel on the prophecy register $V$ with strength
$\lambda \in [0,1]$,
\[
\rho_{\mathrm{rel}}(\sigma)
= (1-\lambda)\,\sigma + \lambda\,\frac{\id_V}{2}\otimes \Tr_V(\sigma).
\]
Concretely this replaces the $V$-marginal with maximal mixedness with weight $\lambda$.

\subsection{Agent Response Map}

We model the agent as applying a conditional unitary on $X$ based on the measurement outcome:
\[
U_0 = \id_X,\qquad U_1 = X\ (\text{Pauli}-X).
\]
The agent map in instrument form is
\[
\pi(\rho) = \sum_{i=0}^1 (U_i\otimes\id_V)\,(\id_X\otimes M_i)\,\rho\,(\id_X\otimes M_i^\dagger)\,(U_i^\dagger\otimes\id_V).
\]

\subsection{Full Operator $\Phi^{\mathcal{Q}}$ and One-Step Evaluation}

With the above choices the full operator is
\[
\Phi^{\mathcal{Q}}(\rho) = \mathcal{E}\Big(
        \sum_i (U_i\otimes\id)\ \rho_{\mathrm{rel}}\big( (\id\otimes M_i)\rho(\id\otimes M_i^\dagger)\big)\ (U_i^\dagger\otimes\id)
    \Big).
\]

We evaluate $\Phi^{\mathcal{Q}}(\rho_0)$ numerically for the chosen parameter set
$(p,\gamma,\alpha,\lambda)=(0.7,0.2,0.6,0.3)$. The resulting density matrix (rounded to six
decimal places) is:

\[
\rho^{(1)}=\Phi^{\mathcal{Q}}(\rho_0) \approx
\begin{pmatrix}
0.537600 & 0       & 0       & 0.114766\\[4pt]
0        & 0.302400& 0.114766& 0       \\[4pt]
0        & 0.114766& 0.102400& 0       \\[4pt]
0.114766 & 0       & 0       & 0.057600
\end{pmatrix}.
\]

\subsection{Crash Probability and Its Reduction}

Define the crash projector on the latent register
\[
P_{\mathrm{crash}} = \ket{1}\!\bra{1}_X \otimes \id_V.
\]
The initial crash probability is
\[
P_{\mathrm{crash}}(\rho_0) = \Tr\!\big(P_{\mathrm{crash}}\rho_0\big) = 0.300000,
\]
while after one iteration:
\[
P_{\mathrm{crash}}(\rho^{(1)}) = \Tr\!\big(P_{\mathrm{crash}}\rho^{(1)}\big) \approx 0.160000.
\]
This demonstrates that, for this parameter choice, the combination of measurement, regulated
release, and agent action reduces the crash probability substantially in one step.

\subsection{Mutual Information (Leakage) Evaluation}

We compute the mutual information between the latent register $X$ and the regulator-released
prophecy $V_{\mathrm{rel}}$ after the measurement+regulator step. For the intermediate released
state $\rho_{X V_{\mathrm{rel}}}$ obtained during the pipeline, the von Neumann entropies yield:

\[
\begin{aligned}
    S(\rho_X) &\approx 0.881291\ \text{bits},\\
    S(\rho_{V_{\mathrm{rel}}}) &\approx 0.942683\ \text{bits},\\
    S(\rho_{X V_{\mathrm{rel}}}) &\approx 1.233835\ \text{bits},
\end{aligned}
\qquad\Longrightarrow\qquad
I(X;V_{\mathrm{rel}}) \approx 0.590139\ \text{bits}.
\]

Thus the regulator-released signal carries roughly $0.59$ bits of information about the latent
crash register under these parameters.

\subsection{Contraction Test (Numerical)}

To test contraction empirically, we compare two nearby initial states and observe the trace-norm
reduction after one application of $\Phi^{\mathcal{Q}}$.

Let $\rho_0$ correspond to $p=0.700000$ and $\widetilde{\rho}_0$ correspond to
$p'=0.701000$ (a small perturbation). The trace-norm distance between these inputs is
\[
\|\rho_0 - \widetilde{\rho}_0\|_1 \approx 0.0021832203.
\]
After one step we obtain outputs $\rho^{(1)}$ and $\widetilde{\rho}^{(1)}$ and compute
\[
\|\rho^{(1)} - \widetilde{\rho}^{(1)}\|_1 \approx 0.0014670676 < 0.0021832203.
\]
The output distance is smaller than the input distance, consistent with contractive behavior for
this perturbation and parameter choice. Repeated iteration shows convergence towards a fixed
point in this parameter regime.

\subsection{Grid Search for Ethical Regulator Parameters}

We perform a brute-force grid search on $(\lambda,\alpha)\in\{0,0.1,\dots,1.0\}^2$ to empirically
find parameter choices minimizing the single-step harm measured after convergence to a fixed point.
We use the harm functional
\[
H(\rho) = c_1 P_{\mathrm{crash}}(\rho) + c_2 \|\rho - M(\rho)\|_1,
\]
with $c_1=1$, $c_2=0.5$. For each grid cell we iterate $\Phi^{\mathcal{Q}}_{\lambda,\alpha}$ until
convergence (or a fixed maximum number of iterations) and evaluate $H$ at the approximate fixed point.

The grid search returns the optimal parameter pair at
\[
(\lambda^\ast,\alpha^\ast) = (0.0,\,1.0)
\]
with corresponding values
\[
H^\ast \approx 0.13333313,\quad P_{\mathrm{crash}}(\rho^\ast)\approx 0.13333333,\quad
\|\rho^\ast - M(\rho^\ast)\|_1\approx 0.
\]
This indicates that, in this toy model, an un-noised (zero $\lambda$) but sharp measurement
(high $\alpha$) minimizes harm when leakage is not constrained. However, the optimized regime
must be checked against leakage constraints: for $(\lambda^\ast,\alpha^\ast)$ the mutual information
at the fixed point is
\[
I(X;V_{\mathrm{rel}}) \approx 0.269187\ \text{bits},
\]
which may or may not satisfy an ethical leakage bound $\epsilon$ specified by a regulator.

\subsection{Algorithmic Pseudocode (Simulation)}

Below is a compact pseudocode description for simulating $\Phi^{\mathcal{Q}}$, finding fixed
points, and performing the regulator grid search.

\begin{verbatim}
Inputs: rho0, K_set (for E), M0,M1, U0,U1, lam_grid, alpha_grid
Function DEP_V(rho, lam): apply depolarizing-on-V channel
Function M_instrument(rho, alpha): compute M0,M1 per alpha; return list of components C_i = (I⊗M_i) rho (I⊗M_i)
Function PI_from_components(components, lam): for each i: comp_rel = DEP_V(component_i, lam); apply U_i on X; sum -> rho_pi
Function Evolve(rho_pi): apply E via Kraus K_j⊗I sum -> rho_next
Function PHI(rho, lam, alpha): components = M_instrument(rho, alpha); rho_pi = PI_from_components(components, lam); return Evolve(rho_pi)

FixedPointIteration:
  rho = rho0
  for t in 1..T_max:
    rho_next = PHI(rho, lam, alpha)
    if trace_norm(rho_next - rho) < tol: return rho_next
    rho = rho_next
  return rho

GridSearch:
  best = None
  for lam in lam_grid:
    for alpha in alpha_grid:
      rho_fp = FixedPointIteration(rho0, lam, alpha)
      H = harm(rho_fp)
      if best is None or H < best.H:
         best = (lam,alpha,H,rho_fp)
  return best
\end{verbatim}

\subsection{Implementation Notes and Reproducibility}

\begin{itemize}
    \item \textbf{Library suggestions:} Use QuTiP or Qiskit for quantum objects and channels. The
    algorithm above translates directly into NumPy/Scipy linear algebra using Kronecker products.
    \item \textbf{Numeric stability:} use double precision and ensure positivity by symmetrization
    $\rho \leftarrow (\rho + \rho^\dagger)/2$ when small numerical asymmetries occur.
    \item \textbf{Convergence criteria:} set $\mathrm{tol}=10^{-8}$ for fixed-point convergence
    in research-grade experiments; looser tolerances suffice for exploratory work.
    \item \textbf{Performance:} for small Hilbert spaces (2--4 qubits) direct matrices are fine;
    for larger spaces use sparse representations and exploit Kronecker structure.
\end{itemize}

\subsection{Observations and Discussion}

The worked example demonstrates several important phenomena:

\begin{enumerate}
    \item \emph{Reflexive mitigation:} agent actions conditioned on measured prophecy can
    materially reduce crash probability (from $0.30$ to $0.16$ in one step for our example).
    \item \emph{Contraction in practice:} with moderate noise and measurement parameters the
    operator exhibits empirically contractive behavior, implying convergence to a reflexive
    equilibrium.
    \item \emph{Ethical trade-offs:} minimization of harm in this toy model favored sharp
    measurements (high $\alpha$) and no regulator noise, but this increases information leakage;
    thus a regulator bound on $I(X;V_{\mathrm{rel}})$ can force more conservative parameter choices.
    \item \emph{Computability caveats:} although the toy model is decidable and convergent,
    Section~6 shows regimes where embedded computation renders prophecy undecidable — such
    regimes must be detected and avoided in practice (safety abort rules).
\end{enumerate}

\subsection{Deliverables and Next Steps}

From this implementation we recommend the following immediate deliverables:
\begin{itemize}
    \item A reproducible notebook (Jupyter) implementing the pseudocode and parameter grid.
    \item Visualization dashboards showing convergence, P\_crash vs.\ iteration, and leakage vs. harm.
    \item An automated safety detector that flags non-contracting behavior (e.g., rising trace distances
    over iterations) to prevent entry into undecidable regimes.
    \item A short experimental study exploring the Pareto frontier between leakage $\epsilon$ and
    achievable harm $H(\rho^\ast)$.
\end{itemize}

This completes the worked example and provides a concrete, reproducible path to implementing,
testing, and iterating the Quantum Reflexive Control Engine in simulation and (when feasible)
on near-term quantum hardware emulators.
